Crime in Austin Texas: Exploration, Classification, and Trends

### 2024-08-07 products/manuscript/Manuscript.qmd
last commit for final project

[Last year](https://www.fox7austin.com/news/austin-named-city-with-15th-biggest-homicide-rate-problem-study), Austin was ranked among the top US cities with a problem in homicide rate [@fox2023].
Earlier [this year](https://www.kvue.com/article/news/local/austin-crime-rate-drops-2024-2020-comparison/269-b62042b7-a6f2-4224-87ff-ede655a37912), the claim was made that crime has dropped to a new low since 2020 [@fox2024].
The intention of this analysis will be to investigate both claims, and gain a better understanding of Austin's criminal activity.
While these claims will be in mind throughout the steps of the analysis, I intend to approach it from more of an exploratory standpoint, so not to only go out and investigate the claims but really any insights that can be gleaned around Austin criminal activity and how it has changed over time.
What the analysis will not cover would be things like *causes* of the changes in crime, at least not directly.
I may identify shifts that could inform the cause, but the goal will not be to drill down into every potential socioeconomic factor and pick the primary drivers, instead simply to observe what crime in the city of Austin looks like and what we might be able to expect going forward.
Aside from resources outlined in our course resources, I will also draw from this [book](https://otexts.com/fpp3/) [@hyndman2021].
It's not one I have used before but it had been recommended to me previously and I would like to learn more about time series and forecasting in general.

@@

My primary data source will come from [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data).
The data is updated weekly by the Austin Police Department, and each record in the dataset represents an Incident Report, with the highest offense within an incident taking precedence in things like the description and categorization Each Incident can have other offense tied to it, however since each record is a unique incident then only the aforementioned Higehst Offense is the one represented (NOTE: At the time of this writing, 7/31/2024, the dataset has been taken down to be replaced with one that aligns more closely with the FBI National Incident Based Reporting System. The datasets are not one-to-one, so reproducibility would require more than a lift and shift, however some of the methods could still be employed).\
The raw data is represented by several categorical, location, and time-based variables, many of which have missing values, or are formatted incorrectly for the data type, so they will need to be cleaned or recoded.
After cleaning the data (done in a separate file found in this repo) we can take a look a more meaningful look.

>>>>

[Last year](https://www.fox7austin.com/news/austin-named-city-with-15th-biggest-homicide-rate-problem-study), Austin was ranked among the top US cities with a problem in homicide rate [@fox2023].
Earlier [this year](https://www.kvue.com/article/news/local/austin-crime-rate-drops-2024-2020-comparison/269-b62042b7-a6f2-4224-87ff-ede655a37912), the claim was made that crime has dropped to a new low since 2020 [@fox2024].
The intention of this analysis will be to investigate both claims, and gain a better understanding of Austin's criminal activity.
While these claims will be in mind throughout the steps of the analysis, I intend to approach it from more of an exploratory standpoint, so not to only go out and investigate the claims but really any insights that can be gleaned around Austin criminal activity and how it has changed over time.
What the analysis will not cover would be things like *causes* of the changes in crime, at least not directly.
I may identify shifts that could inform the cause, but the goal will not be to drill down into every potential socioeconomic factor and pick the primary drivers, instead simply to observe what crime in the city of Austin looks like and what we might be able to expect going forward.
Aside from resources outlined in our course resources, I will also draw from this [book](https://otexts.com/fpp3/) [@hyndman2021].
The final output of this project, the manuscript you are reading, is rendered in my github analytics [portfolio](https://whygit10.github.io/williamhyltin-P2-portfolio/). The original repo dedicated to just this project can be found [here](https://github.com/whygit10/HYLTIN-PII-project).

@@

My primary data source will come from [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data).
The data is updated weekly by the Austin Police Department, and each record in the dataset represents an Incident Report, with the highest offense within an incident taking precedence in things like the description and categorization Each Incident can have other offense tied to it, however since each record is a unique incident then only the aforementioned Higehst Offense is the one represented (NOTE: At the time of this writing, 7/31/2024, the dataset has been taken down to be [replaced](https://data.austintexas.gov/Public-Safety/NIBRS-Group-A-Offense-Crimes/i7fg-wrk5/about_data) with one that aligns more closely with the FBI National Incident Based Reporting System. The datasets are not one-to-one, so reproducibility would require more than a lift and shift, however some of the methods could still be employed).\
The raw data is represented by several categorical, location, and time-based variables, many of which have missing values, or are formatted incorrectly for the data type, so they will need to be cleaned or recoded.
After cleaning the data (done in a separate file found in this repo) we can take a look a more meaningful look.
